Sperm should evolve to make female meiosis fair

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
incorporated coops comments

In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype \cite[e.g. ][ and see Model 1 in the Appendix for more details]{Ubeda2004}. %The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
	%YB - MOVED THIS TO THE APPENDIX
          To depict a case
of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm}A and Model 2 in the Supplementary Worksheet). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

For comparative purposes, we briefly present the standard drive model %as Model 1 in the Appendix, 
%equations \ref{eqn.A1}-\ref{eqn.A3}, 
	%and refer readers to the previous literature on this model 
	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
Assuming the driving allele is deleterious in both sexes, but fully recessive 
	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
will be maintained as a protected polymorphism \citep{Prout1973}. 
The parameter space where the allele can invade but not fix is shown in white
        in Figure \ref{Eggsperm}B. %, and it can fix in the colored regions. 
When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
	drive suppressors, corresponding well to empirical examples of
        female meiotic drive \citep[reviewed in ][]{Burt2006}.

In contrast to a traditional driver, which drives but pays effectively
	no fitness cost when rare, 
a self-promoting driver specifically creates low fitness drive homozygotes 
	by uniting driving female gametes with sperm enabling that drive.
        % zygotes when it drives. in heterozygotes. 
%, and therefore there is a drive-associated fitness cost even when rare. 
It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
The conditions allowing the invasion of a self-promoting driver
 	are consequently far more restrictive than those for a standard meiotic driver.
When rare, a fully recessive, self-promoting driver can only invade when $s$ 
	%, the selection against the drive homozygote, 
	is less than $(2 d - 1)/(4 d)$. 
This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).

@@

Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.  
Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
	% fixes when $s$ is less than $d_{hom} - 1/2$.
However, when the distorting effect of genotype displays strong dominance in its
	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
	there is a a narrow sliver of parameter space sustains a polymorphism
	when the cost of the drive is recessive
	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
%Presumably this zone of polymorphism comes about because positive frequency 
%	dependence diminishes with the dominance of the male effect driver. 
%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
While mathematically interesting, it does not seem particularly realistic to think that the
	effect of the drive allele would be dominant in its action through the
	male genotype, while the cost would be recessive. 
Therefore, although Model 3 can sustain a polymorphism, 
	the lack of biological reality underlying the small portion of parameter values 
	required for this polymorphism make us doubt it general applicability. 
%Thus we regard this case as interesting but not very general.

@@

Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, 
	the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

>>>>

In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype \cite[e.g. ][ and see Model 1 in the Appendix for more details]{Ubeda2004}. %The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
	%YB - MOVED THIS TO THE APPENDIX
          To depict a case
of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm
        carrying that allele (see Figure \ref{Eggsperm}A and Model 2
        in the Appendix and Supplementary Worksheet). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude our single locus results with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

For comparative purposes, we first briefly present the standard drive model %as Model 1 in the Appendix, 
%equations \ref{eqn.A1}-\ref{eqn.A3}, 
	%and refer readers to the previous literature on this model 
	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
Assuming the driving allele is deleterious in both sexes, but fully recessive 
	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B), a driver cannot fix, and
will be maintained as a protected polymorphism \citep{Prout1973}. 
The parameter space where the allele can invade but not fix is shown in white
        in Figure \ref{Eggsperm}B. %, and it can fix in the colored regions. 
When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
	drive suppressors, corresponding well to empirical examples of
        female meiotic drive \citep[reviewed in ][]{Burt2006}.

In contrast to a traditional driver, which drives but pays effectively
	no fitness cost when rare, 
a self-promoting driver specifically creates low fitness drive homozygotes 
	by uniting driving female gametes with sperm enabling that drive.
        % zygotes when it drives. in heterozygotes. 
%, and therefore there is a drive-associated fitness cost even when rare. 
It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
The conditions allowing the invasion of a self-promoting driver
 	are consequently far more restrictive than those for a standard meiotic driver.
When rare, a fully recessive, self-promoting driver can only invade when $s$ 
	%, the selection against the drive homozygote, 
	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B). 
This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).

@@

Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
This paternal-genotype dependent model (Model 3 in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.  
Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
	% fixes when $s$ is less than $d_{hom} - 1/2$.
However, when the distorting effect of genotype displays strong dominance in its
	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
	there is a a narrow sliver of parameter space sustains a polymorphism
	when the cost of the drive is recessive
	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
%Presumably this zone of polymorphism comes about because positive frequency 
%	dependence diminishes with the dominance of the male effect driver. 
%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
While mathematically interesting, it does not seem particularly realistic to think that the
	effect of the drive allele would be dominant in its action through the
	male genotype, while the cost would be recessive. 
Therefore, although Model 3 can sustain a polymorphism, 
	the lack of biological reality underlying the small portion of parameter values 
	required for this polymorphism make us doubt it general applicability. 
%Thus we regard this case as interesting but not very general.

@@

Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles, Figure \ref{Eggsperm_3_allele_cartoon}A1). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
        Model 4 in the Appendix and Supplementary Worksheet). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, 
	the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$ (Figure
          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Equations \ref{A-} and
        \ref{unlinked}).  %%DELETED FIG REF HERE.
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

### 2014-05-20 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
incorporated coops comments

\section{Results}
\subsection{Invasion of an allele that promotes its own drive}
In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004} and see Model 1 in the Appendix for more details). 
To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Supplementary Worksheet). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

For comparative purposes, we briefly present the standard drive model 
	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
Assuming the driving allele is deleterious in both sexes, but fully recessive 
	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
	it always invades because, when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
will be maintained as a protected polymorphism \cite{Prout1973}. 
The parameter space where the allele can invade but not fix is shown in white
        in Figure 1B. . 
When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
	drive suppressors, corresponding well to empirical examples of
        female meiotic drive (reviewed in \cite{Burt2006}).

In contrast to a traditional driver, which drives but pays effectively
	no fitness cost when rare, 
a self-promoting driver specifically creates low fitness drive homozygotes 
	by uniting driving female gametes with sperm enabling that drive.
It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
The conditions allowing the invasion of a self-promoting driver
 	are consequently far more restrictive than those for a standard meiotic driver.
When rare, a fully recessive, self-promoting driver can only invade when $s$ 
	is less than $(2 d - 1)/(4 d)$. 
This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
	closely matches results obtained by exact numerical iteration (Figure 1B).

@@

Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm (e.g. \cite{Joseph2004}) 
	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.  
Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
However, when the distorting effect of genotype displays strong dominance in its
	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
	there is a a narrow sliver of parameter space sustains a polymorphism
	when the cost of the drive is recessive
	(see Figure S3, and Equation \ref{polymale}).  
While mathematically interesting, it does not seem particularly realistic to think that the
	effect of the drive allele would be dominant in its action through the
	male genotype, while the cost would be recessive. 
Therefore, although Model 3 can sustain a polymorphism, 
	the lack of biological reality underlying the small portion of parameter values 
	required for this polymorphism make us doubt it general applicability.

@@

\subsection{Two locus models of sperm-dependent female drive}
Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2, and Model 4 in the Supplementary Worksheet). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Mathematical Supplement).

While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     2, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure 2 and Equations \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure 2). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems (e.g. \cite{Brandvain2012}).  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

>>>>

\section{Results}
\subsection{Invasion of an allele that promotes its own drive}
In the standard single-locus, biallelic model of female meiotic drive,
the driving allele is transmitted to the egg in heterozygotes with
probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004} and see Model 1 in the Appendix for more details). 
To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Supplementary Worksheet). 
We then identify the conditions allowing for the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
We conclude our single locus results with an analysis of a related model (Model 3) - in which
drivers influence their transmission in females via paternal genotype,
rather than sperm haplotype.

For comparative purposes, we first briefly present the standard drive model 
	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
Assuming the driving allele is deleterious in both sexes, but fully recessive 
	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
	it always invades because, when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure 1B), a driver cannot fix, and
will be maintained as a protected polymorphism \cite{Prout1973}. 
The parameter space where the allele can invade but not fix is shown in white
        in Figure 1B. . 
When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
	drive suppressors, corresponding well to empirical examples of
        female meiotic drive (reviewed in \cite{Burt2006}).

In contrast to a traditional driver, which drives but pays effectively
	no fitness cost when rare, 
a self-promoting driver specifically creates low fitness drive homozygotes 
	by uniting driving female gametes with sperm enabling that drive.
It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
The conditions allowing the invasion of a self-promoting driver
 	are consequently far more restrictive than those for a standard meiotic driver.
When rare, a fully recessive, self-promoting driver can only invade when $s$ 
	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure 1B. 
This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
	closely matches results obtained by exact numerical iteration (Figure 1B).

@@

Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm (e.g. \cite{Joseph2004}) 
	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
This paternal-genotype dependent model (Model 3  in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.  
Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
However, when the distorting effect of genotype displays strong dominance in its
	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
	there is a a narrow sliver of parameter space sustains a polymorphism
	when the cost of the drive is recessive
	(see Figure S3, and Equation \ref{polymale}).  
While mathematically interesting, it does not seem particularly realistic to think that the
	effect of the drive allele would be dominant in its action through the
	male genotype, while the cost would be recessive. 
Therefore, although Model 3 can sustain a polymorphism, 
	the lack of biological reality underlying the small portion of parameter values 
	required for this polymorphism make us doubt it general applicability.

@@

\subsection{Two locus models of sperm-dependent female drive}
Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
	with alternative models of two loci - one a female driver, 
	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
In this model, a female meiotic driver with no sperm-dependency 
	initially reaches drive-viability equilibrium (with two alleles
	$A$ and $B$ are the ancestral non-driver and driver alleles, Figure 2A1). 
Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.

@@

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Supplementary Worksheet). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Mathematical Supplement).

While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, (Figure 2A3, Model 4 in Appendix). 
 This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     2, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Equations \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure 2B2). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems (e.g. \cite{Brandvain2012}).  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.
